---
title: "Data Science Module 3 Exercises"
author: "Brady Lamson"
date: "2/16/2022"
output:
    pdf_document:
        dev: png
        includes:
            in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    out.width = "70%"
    )
library(dplyr)
library(ggplot2)
library(nycflights13)
```

# 4: Data Wrangling
## 4.2: 
Extracting Columns with select()

### Exercise 1
a) This will pull only the years and days from the flights data set.

b) This pulls all the columns between year and day, inclusive.

c) This pulls all columns *except* year and day.

***

### Exercise 2:
a) This will pull all column names starting with "sched", of which there are 2 columns in this data set. (sched_dep_time and sched_arr_time)

b) Same as (a) but with columns starting with "arr" (arr_time and arr_delay).

c) Pulls all columns starting either with "dep_" or "arr_".

***

\pagebreak
## 4.4: Filtering Rows with filter()
### Exercise 3:

```{r, results='hide'}
# A
dplyr::filter(flights, arr_delay >= 120)
```

```{r, results='hide'}
# B 
flights %>%
    dplyr::filter(dest %in% c("IAH", "HOU"))
```

```{r, results='hide'}
# C
flights %>%
    dplyr::filter(carrier %in% c("UA", "AA", "DL"))
```

```{r, results='hide'}
# D
flights %>%
    dplyr::filter(month %in% 7:9)
```

```{r, results='hide'}
# E
flights %>%
    dplyr::filter(dep_time %in% 0:600)
```

```{r, results='hide'}
# F
flights %>%
    dplyr::filter(
        carrier == "UA", 
        month == 7, 
        arr_delay >= 120
    )
```

***
\pagebreak
## 4.5: Arranging Rows with arrange()
### Exercise 4:

```{r, results='hide'}
# A
flights %>%
    dplyr::arrange(dep_delay, arr_delay)
```

```{r, results='hide'}
# B
flights %>%
    dplyr::arrange(
        dplyr::desc(dep_delay),
        dplyr::desc(arr_delay)
    )
```

```{r, results='hide'}
# C
flights %>%
    dplyr::arrange(dep_time)
```

```{r, results='hide'}
# D
flights %>%
    dplyr::arrange(dplyr::desc(dep_time))
```

```{r, results='hide'}
# E
flights %>%
    dplyr::arrange(distance)
```

```{r, results='hide'}
# F
flights %>%
    dplyr::arrange(dplyr::desc(distance))
```

***
\pagebreak
### Exercise 5:

```{r}
x <- data.frame(x1 = c(2, 1, NA, 8, 7, 5, 4),
x2 = c("a", NA, "c", "d", "c", "a", "d"),
stringsAsFactors = FALSE)
```


The following code will sort the data with NAs at the top of the *first* column.

```{r}
x %>%
    dplyr::arrange(is.na(x1))
```


The following code will sort the data with NAs at the bottom of the *first* column.

```{r}
x %>%
    dplyr::arrange(
        dplyr::desc(is.na(x1))
    )
```

***
\pagebreak
## 4.6: Create New Variables (Columns) with mutate()
### Exercise 6:

```{r, results='hide'}
flights %>%
    dplyr::mutate(travel_time = arr_time - dep_time)
```

The arrival and departure time use totally different units than air time. The former are units of time, so 600 is 6:00am, whereas with air time 600 would be 600 minutes. They aren't equivalent. 

***
### Exercise 7:
```{r}
flights_small <- select(
    .data = flights, 
    year:day, 
    ends_with("delay"), 
    distance,
    air_time)
```

```{r}
dplyr::mutate(
    .data = flights_small,
    gain = dep_delay - arr_delay,
    hours = air_time / 60,
    gain_per_hour = gain / hours
    )
```

The variable gain_per_hour does get computed! 

***
\pagebreak
## 4.7: Renaming Variables (Columns) with rename()
### Exercise 8:

```{r}
z <- data.frame(
    z1 = c(5, 4, 3),
    z2 = c("a", "c", "b"),
    z3 = c(14, 22, 13))
```

```{r}
z %>%
    dplyr::rename(
        new_z1 = z1,
        new_z2 = z2,
        new_z3 = z3
    )
```

***
\pagebreak
## 4.8: Summarize Data with summarize()

```{r}
not_cancelled <- filter(.data = flights, !is.na(dep_delay), !is.na(arr_delay))
```

### Exercise 9:

```{r}
#A

not_cancelled %>%
    dplyr::summarise(
        median_dep_delay = median(dep_delay),
        median_arr_delay = median(arr_delay)
    )
```

```{r}
# B
not_cancelled %>%
    dplyr::summarise(
        max_dep_delay = max(dep_delay),
        max_arr_delay = max(arr_delay)
    )
```

```{r}
not_cancelled %>%
    dplyr::summarise(
        shortest_dep_delay = min(dep_delay),
        shortest_arr_delay = min(arr_delay)
    )
```

***
\pagebreak
### Exercise 10:
a) The following code will give you the total count of rows in the data set. 

```{r}
summarize(.data = not_cancelled,
          total_flights = n())
```


b) This counts the *number* of flights that have a delay of an hour or more. 

```{r}
summarize(.data = not_cancelled,
          hour_arr_delay_total = sum(arr_delay > 60))
```


c) This returns the ratio of flights with a delay of an hour or more compared to the total number of flights in the data set.

```{r}
summarize(.data = not_cancelled,
          hour_arr_delay_proportion = sum(arr_delay > 60) / n())
```

***
\pagebreak
## 4.9: Applying summarize() to Groups using group_by()
### Exercise 11:

```{r}
# A

by_dest <- group_by(.data = flights, dest)

delay_by_dest <- summarize(
    .data = by_dest, 
    mean_arr_delay = mean(arr_delay, na.rm = TRUE)
    )
```

delay_by_dest gives a breakdown of the average delay by the location of travel.

```{r}
# B
by_dest <- group_by(.data = flights, dest)

delay_dist_by_dest <- summarize(
    .data = by_dest,
    mean_dist = mean(distance, na.rm = TRUE),
    mean_arr_delay = mean(arr_delay, na.rm = TRUE)
)
```

***
\pagebreak
### Exercise 12:
```{r}
resp <- c(23, 11, 14, 16, 19, 26, 24, 29, 31, 28, 34, 25)
trt <- c(rep("Ctrl", 4), rep("TrtA", 4), rep("TrtB", 4))
age <- c(33, 45, 30, 24, 22, 31, 39, 40, 29, 19, 27, 25)
gndr <- c("m", "m", "f", "f", "m", "f", "f", "m", "f", "m", "f", "m")

ExpData <- data.frame(TrtGrp = trt,
                      SubjectGender = gndr,
                      SubjectAge = age,
                      Response = resp, stringsAsFactors = FALSE)
```

```{r}
# A

by_TrtGrp <- group_by(.data = ExpData, TrtGrp)
summarize(.data = by_TrtGrp, Count = n())
```

This command presents the total count of each treatment group. So we have 4 observations in the control group, 4 in treatment A and 4 in treatment B. 

```{r}
# B

dplyr::summarize(.data = by_TrtGrp, mean = base::mean(Response))
```

```{r}
# C

dplyr::summarise(
    by_TrtGrp, 
    mean_response = base::mean(Response), 
    mean_subject_age = base::mean(SubjectAge)
)
```

### Exercise 13:

```{r}
# A

plane_count <- flights %>%
    dplyr::group_by(tailnum) %>%
    dplyr::summarise(count = n()) %>%
    dplyr::arrange(dplyr::desc(count)) %>%
    stats::na.omit()

glue::glue("
    The tail number with the most flights is {plane_count[1,1]}.
")
```

```{r}
# B

dest_count <- flights %>%
    dplyr::group_by(dest) %>%
    dplyr::summarise(count = n()) %>%
    dplyr::arrange(dplyr::desc(count)) %>%
    stats::na.omit()

glue::glue("
    The destination visited the most times was {dest_count[1,1]}.
")
```

***
\pagebreak
## 4.10: Chaining Together Actions Using the Pipe Operator %>%
### Exercise 14:
```{r}
# A

x <- c(2, 5, 4, 3, 7, 9)
# x %>% mean()
```

The above command takes the mean of the vector x. It's equivalent to mean(x)

```{r}
# B

# x %>% mean() %>% sqrt() %>% round(digits = 2)
```

Take the mean of x, then the square root, and then round that output to 2 digits. 

c) Rewrite the command below. 

```{r}
mean_x <- mean(x)
sqrt_mean_x <- sqrt(mean_x)
round_sqrt_mean_x <- round(sqrt_mean_x, digits = 2)
```

```{r}
x %>%
    mean() %>%
    sqrt() %>%
    round(digits = 2)
```
d) Rewrite the command below.

```{r}
round(sqrt(mean(x)), digits = 2)
```

```{r}
x %>%
    mean() %>%
    sqrt() %>%
    round(digits = 2)
```

***
\pagebreak
### Exercise 15

Rewrite the given commands using pipes. 

```{r}
# A

flights %>%
    dplyr::select(arr_delay) %>%
    head()
```

```{r}
# B

flights %>%
    dplyr::select(dest, arr_delay) %>%
    head()
```

```{r}
# C

flights %>%
    dplyr::select(dest, arr_delay) %>%
    dplyr::filter(dest %in% c("SEA", "DEN")) %>%
    head()
```

```{r}
# D

flights %>%
    dplyr::select(dest, arr_delay) %>%
    dplyr::filter(dest %in% c("SEA", "DEN")) %>%
    dplyr::group_by(dest) %>%
    dplyr::summarise(mean_arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
    head()
```

***
### Exercise 16:
Rewrite the following command using pipes.

```{r}
flights %>%
    dplyr::filter(dest == "DEN") %>%
    dplyr::summarise(
        mean_dep_delay = mean(dep_delay, na.rm = TRUE),
        mean_arr_delay = mean(arr_delay, na.rm = TRUE)
    )
```

***
\pagebreak
## 4.11: Combining Multiple Data Frames
### Exercise 17:
```{r}
df1 <- data.frame(Respondent_ID = c(1001, 1002, 1003),
                  Q1_Response = c(55, 62, 39))

df2 <- data.frame(Respondent_ID = c(1002, 1003, 1004),
                  Q2_Response = c("yes", "no", "yes"))
```

Guess the results of the following commands. 

a) This will return a data set with *only* ID 1002 and 1003, with Q1 and Q2 responses for each. 

```{r}
# A

inner_join(x = df1, y = df2, by = "Respondent_ID")
```
b) This will give Q2 responses to 1002 and 1003, but 1001's Q2 response will be NA. 

```{r}
# B

left_join(x = df1, y = df2, by = "Respondent_ID")
```

c) This will give a data set with all IDs in either data set. The Q1 and Q2 responses will either be filled in if they exist or left NA if they do not.

```{r}
# C

full_join(x = df1, y = df2, by = "Respondent_ID")
```

\pagebreak
d) I'm guessing it'd default to Respondent ID, because that column is in both data frames. So this would be just like (c).

```{r}
# D

full_join(x = df1, y = df2)
```

e) This would create a single Response column for the joined data frame, this one with both numbered Q1 responses AND yes/no responses from Q2. This seems like a bad idea.

**update:** It actually returned an error. This makes sense in retrospect as the values in a column (a vector) must all be the same type. 

```{r}
# E

df1 <- rename(.data = df1, Response = Q1_Response)
df2 <- rename(.data = df2, Response = Q2_Response)

#full_join(x = df1, y = df2)
```

f) This would either give the same error as last time or do nothing.

**update:** My guess was correct, it gave the same error as in part (e).
```{r}
# F

# inner_join(x = df1, y = df2)
```

***
\pagebreak
### Exercise 18:

```{r}
df1 <- data.frame(Respondent_ID = c(1000, 1001, 1002, 1003, 1004, 1005, 1006),
                  Q1_Response = c(55, 62, 39, 45, 70, 77, 56))

df2 <- data.frame(Respondent_ID = c(1003, 1002, 1000, 1004, 1006, 1001, 1005),
                  Q2_Response = c(12, 17, 23, 24, 19, 30, 20))
```

```{r}
# A

inner_join(x = df1, y = df2, by = "Respondent_ID")
```

```{r}
# B

inner_join(x = df2, y = df1, by = "Respondent_ID")
```

a) Thankfully it's smart enough to pick the values that match, so the responses are allocated accordingly.

b) This shows how the merged data frames are ordered, and that's by the data frame assigned to x. 

***
\pagebreak
### Exercise 19

```{r}
dfX <- data.frame(LastName = c("Smith", "Smith", "Jones", "Smith",
                               "Olsen", "Taylor", "Olsen"),
                  FirstName = c("John", "Kim", "John", "Marge", "Bill",
                                "Bill", "Erin"),
                  Gender = c("M", "F", "M", "F", "M", "M", "F"),
                  ExamScore = c(75, 80, 64, 78, 90, 89, 79))

dfY <- data.frame(LastName = c("Olsen", "Jones", "Taylor", "Smith",
                               "Olsen", "Smith", "Smith"),
                  FirstName = c("Bill", "John", "Bill", "Kim", "Erin",
                                "John", "Marge"),
                  Gender = c("M", "M", "M", "F", "F", "M", "F"),
                  Grade = c("A", "D", "B", "B", "C", "C", "C"))
```

```{r}
# A 

full_join(dfX, dfY, by = c("LastName", "FirstName", "Gender"))
```

```{r}
# B

full_join(x = dfX, y = dfY, by = c("LastName", "FirstName"))
```

This creates two separate gender columns, one from the x data frame and one from the y data frame.

\pagebreak

```{r}
# C

full_join(x = dfX, y = dfY, by = "LastName")
```

This creates an abomination of a data frame, having separate first name and gender columns! It essentially renders the data useless.