---
title: "Module 2 Exercises"
author: "Brady Lamson"
date: "2/7/2022"
output: 
    pdf_document:
        dev: png
        includes:
            in_header: "hw_2/preamble.tex"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    out.width = "70%"
    )
library(dplyr)
library(ggplot2)
```

# 3: Data Visualization (Graphics)
## 3.2: A Taxonomy for Data Graphics
### Exercise 1: 
For each graph indicate:

- The **visual cues** that are used
- The **coordinate system** that's used
- The **scales** that are used
- How **context** is provided

```{r}
ggplot(data = diamonds) +
    geom_point(mapping = aes(x = carat, y = price, color = cut)) +
    ggtitle("Diamond Price vs Weight")
```

a) 

    - **Visual Cues:** *Position* along the x and y axis and *color*.
    
    - **Coordinate System:** *Cartesian*
    
    - **Scale:** *Numerical* for the axes scale *Categorical* for the colors
    
    - **Context:** Legend, x and y-axis labels, title

\pagebreak
```{r}
ggplot(data = diamonds) +
geom_point(mapping = aes(x = carat, y = price)) +
geom_smooth(mapping = aes(x = carat, y = price), se = FALSE) +
ggtitle("Diamond Price vs Weight")
```

b) 

    - **Visual Cues:** *Position* along the x and y axis. *Angle* / *direction* of the line

    - **Coordinate System:** *Cartesian*

    - **Scale:** *Numerical*

    - **Context:** Legend, x and y-axis labels, title

\pagebreak
```{r}
ggplot(data = diamonds) +
geom_histogram(mapping = aes(x = price)) +
ggtitle("Histogram of Diamond Prices")
```

c) 

    - **Visual Cues:** *Position* along the x and y axis, *length* of the histogram bars. *Area* of the histogram bars. 

    - **Coordinate System:** *Cartesian*
    
    - **Scale:** *Numerical*
    
    - **Context:** X and y-axis lables, title.

***
\pagebreak
### Exercise 2:
```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, size = cyl)) +
ggtitle("Highway MPG vs Displacement")
```

a) 

    - **Visual Cues:** *Position* along the x and y axis, *area* of the circles. 

    - **Coordinate System:** *Cartesian*
    
    - **Scale:** *Numerical*
    
    - **Context:** X and y-axis lables, title, legend.
    
```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = cty, size = hwy,
color = cyl, shape = drv)) +
ggtitle("City MPG vs Displacement")
```

b) 

    - **Visual Cues:** *Position* along the x and y axis, *shade*, *shape*, *area*.

    - **Coordinate System:** *Cartesian*
    
    - **Scale:** *Numerical*
    
    - **Context:** Title, x and y-axis, three different legends
    
***
\pagebreak
# 4: A Grammar for Graphics with "ggplot2"
## 4.1: Introduction
### Exercise 3:
```{r, out.width="30%"}
ggplot(data = mpg)
```

This outputs a blank box, the box that will have more and more information added onto it once more things are specified. 

### Exercise 4:
Guess whether the following commands both make the same scatterplot, then check your answer:

I would guess **yes**. These would be different *if* there was another geom_* used with a different dataset. These, I think, should be functionally equivelant. 
```{r, fig.show="hold", out.width="50%", echo=TRUE}
## Specify data in ggplot():
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy))

## Specify data in geom_*() function:
ggplot() +
geom_point(data = mpg, mapping = aes(x = displ, y = hwy))
```

***
\pagebreak
### Exercise 5:

I would make the same guess that **yes**, these are equivalent expressions. Not enough is really going on to impact the graph. 
```{r, fig.show="hold", out.width="50%", echo=TRUE}
## Specify aesthetics in geom_*() function:
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy))
## Specify aesthetics in ggplot():
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point()
```

***
\pagebreak
### Exercise 6

a) Guess what the **ggtitle(), xlab()**, and **ylab()** commands do to the scatterplot below and to the left. Then check your answers.

    - ggtitle will put a title at the top with the text "Highway MPG vs Displacement"
    
    - xlab and ylab will put text labels on the x-axis and y-axis respectively. 

b) Guess what the **labs()** command does to the scatterplot below and to the right. Then check your answer.

    - I'm not sure actually. col = drv means we have some categorical color usage here and my gut instinct says the only way a label would make sense here is with a legend. Let's go with that then! This will create a legend with the text label above it saying "Drivetrain". 

```{r, fig.show="hold", out.width="50%", echo=TRUE}
# A
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy)) +
ggtitle(label = "Highway MPG vs Displacement") +
xlab(label = "Displacement") +
ylab(label = "Highway MPG")

# B
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
labs(color = "Drivetrain")
```

***
\pagebreak
### Exercise 7:
a) Make a scatterplot of hwy (on the y-axis) versus cyl (x-axis). Report your R commands.
```{r}
ggplot(data = mpg, aes(x = hwy, y = cyl)) +
    geom_point() +
    labs(title = "MPG Dataset: Highway Miles per Gallon vs. # of Cylinders")
```

b) Reproduce the scatterplot of Part a, but now add a second layer to the plot using **geom_smooth()**. Report your R command(s).
```{r}
ggplot(data = mpg, aes(x = hwy, y = cyl)) +
    geom_point() +
    geom_smooth() +
    labs(title = "MPG Dataset: Highway Miles per Gallon vs. # of Cylinders")
```

c) Make a scatterplot of class (y-axis) versus drv (x-axis)? What happens? Why is the plot not useful?
```{r}
ggplot(data = mpg, aes(x = drv, y = class)) +
    geom_point()
```

This plot is trying to use Cartesian coordinates for two different categorical values. The plot just doesn't make any sense as none of the visual cues carry any relevant information. 

***
\pagebreak
## 4.2: More on Aesthetic Mappings
