---
title: "Homework 10"
author: "Brady Lamson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(mdsr)
library(RMySQL)
```

```{r}
# Open a connection to the airlines database:
db_conn <- dbConnect_scidb("airlines")
```

First let's see what is even in flights for future reference. 

```{r}
dbGetQuery(conn = db_conn,
           statement = "DESCRIBE flights;"
)
```

***
\pagebreak
# Problem 2

```{r}
dbGetQuery(conn = db_conn,
           statement = "SELECT DISTINCT year
                        FROM flights;"
)
```

***
# Exercise 3

```{r}
dbGetQuery(conn = db_conn,
           statement = "SELECT COUNT(*) AS number_of_flights
                        FROM flights
                        WHERE dest = 'DFW' AND year = 2010 AND month = 5 AND day = 14;"
)
```

***
# Problem 5

```{r}
dbGetQuery(conn = db_conn,
           statement = "SELECT dest, COUNT(*) AS frequency
                        FROM flights
                        WHERE origin = 'ORD' AND year = 2010
                        GROUP BY dest
                        ORDER BY frequency DESC
                        LIMIT 5;"
)
```

***
# Problem 6

```{r}
dbGetQuery(conn = db_conn,
           statement = "SELECT dest, AVG(arr_delay) AS avg_arr_delay
                        FROM flights
                        WHERE year = 2010
                        GROUP BY dest
                        ORDER BY avg_arr_delay DESC
                        LIMIT 5;"
)
```

***
# Problem 8

```{r}
dbGetQuery(conn = db_conn,
           statement = "SELECT carrier, flight
                        FROM flights
                        WHERE year = 1990 AND day = 26 AND month = 9 AND origin = 'LAX' AND dest = 'JFK';"
)
```

This returns 0 flights. This is both surprising and unsurprising. Unsurprising because our previous `SELECT DISTINCT year` query returned only flights from 2010 to 2017, but this is surprising because I don't know why this is a problem being asked in the first place. 

```{r, eval=FALSE, echo=FALSE}
# TESTING W/ DR. GREVSTAD ----------------------------------

# Open a connection to the airlines database:
db_conn <- dbConnect_scidb("airlines")

# Equivalent connection, only have it here for testing purposes.
db_conn <- dbConnect(drv = MySQL(),
                    dbname = "airlines",
                    host = "mdsr.cdc7tgkkqd0n.us-east-1.rds.amazonaws.com",
                    user = "mdsr_public",
                    password = "ImhsmflMDSwR")

dbGetQuery(conn = db_conn,
           statement = "SELECT DISTINCT year
                        FROM flights;"
)

dbGetQuery(conn = db_conn,
           statement = "SELECT carrier, flight
                        FROM flights
                        WHERE year = 1990;"
)
```