---
title: "Data Science Module 4 Exercises"
author: "Brady Lamson"
date: "2/23/2022"
output: 
    pdf_document:
            dev: png
            includes:
                in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(rvest)
```

# 6: Tidy Data
## 6.2: Using pivot_longer() and pivot_wider()
### Exercise 1
Write a command using pivot_longer() that converts the given data frame to narrow format. Name the columns Grp and Y. 

```{r}
xWide <- data.frame(
    GrpA = c(1, 4, 2, 3),
    GrpB = c(7, 5, 8, 6),
    GrpC = c(9, 9, 8, 7))

xWide %>%
    tidyr::pivot_longer(
        cols = c("GrpA", "GrpB", "GrpC"),
        names_to = "Grp",
        values_to = "Y"
    )
```

***
\pagebreak
### Exercise 2
Here are data from a study in which a variable Y was recorded on each of five subjects before and after an intervention:

```{r}
xNarrow <- data.frame(Subject = c(1:5, 1:5),
                      Period = c("Before", "Before", "Before", "Before",
                                 "Before", "After", "After", "After",
                                 "After", "After"),
                      Y = c(22, 45, 32, 45, 30, 60, 44, 24, 56, 59),
                      stringsAsFactors = FALSE)
```

a) Write a command involving pivot_wider() that converts xNarrow to a **wide** format. 

```{r}
xNarrow %>%
    tidyr::pivot_wider(
        names_from = Period,
        values_from = Y
    )
```

b)

```{r}
data.frame(Period = c("Before", "Before", "Before",
                      "Before", "Before", "After", "After",
                      "After", "After", "After"),
           Y = c(22, 45, 32, 45, 30, 60, 44, 24, 56, 59),
           stringsAsFactors = FALSE) %>%
    tidyr::pivot_wider(
        names_from = Period,
        values_from = Y
    )
```


### Exercise 3
Write a command involving pivot_longer() and the ”helper” function num_range() that converts xWide to narrow format.

```{r}
data.frame(Subject = c(1001, 1002, 1003),
                    t1 = c(22, 45, 32),
                    t2 = c(45, 30, 60),
                    t3 = c(44, 24, 56),
                    t4 = c(55, 27, 53)
                    ) %>%
    tidyr::pivot_longer(
        cols = num_range("t", 1:4),
        names_to = "Time",
        values_to = "Y"
    )
```

### Exercise 4
```{r}
xWide <- data.frame(Subject = c(1001, 1002, 1003),
                    Gender = c("m", "f", "f"),
                    t1 = c(22, 45, 32),
                    t2 = c(45, 30, 60),
                    t3 = c(44, 24, 56),
                    t4 = c(55, 27, 53))
```

What happens to the gender column when you convert xWide to narrow format? Try the below code.

```{r}
xNarrow <- tidyr::pivot_longer(data = xWide,
                        cols = num_range("t", 1:4),
                        names_to = "Time",
                        values_to = "Y")
```

We end up with 8 rows of 'f' because we had two instances of 'f' that were quadrupled.

\pagebreak
## 6.3: Separating and Uniting Columns Using separate() and unite()
### Exercise 5

```{r}
diseases <- data.frame(country = c("Afghanistan", "Afghanistan",
                                   "Brazil", "Brazil", "China", "China"),
                       year = c(1999, 2000, 1999, 2000, 1999, 2000),
                       rate = c("745/19987071", "2666/20595360",
                                "37737/172006362", "80488/174504898",
                                "212258/1272915272", "213766/1280428583"))
```

Write a command that separates the rate column into two columns.

```{r}
diseases %>%
    tidyr::separate(
        col = rate,
        into = c("cases", "population"),
        sep = "/"
    )
```

***
\pagebreak
### Exercise 6

```{r}
year <- c(2017, 2017, 2017, 2017, 2017, 2017, 2017, 2018, 2018, 2018, 2018, 2018,
          2018, 2018)
month <- c(6, 6, 7, 7, 7, 8, 8, 6, 6, 7, 7, 7, 8, 8)
day <- c(4, 18, 2, 16, 30, 13, 27, 3, 17, 1, 15, 29, 12, 26)
phosphate <- c(2.42, 3.50, 1.78, 2.46, 0.66, 1.16, 0.68, 0.90, 1.11, 1.25, 2.28,
               1.36, 0.43, 2.90)
nitrate <- c(3.38, 3.87, 1.28, 3.45, NA, 3.64, 1.88, 6.16, 2.55, 2.98, 3.90, 3.31,
             4.19, 5.35)

river <- data.frame(Year = year,
                    Month = month,
                    Day = day,
                    Phosphate = phosphate,
                    Nitrate = nitrate)
```

```{r}
river %>%
    tidyr::unite(
        col = "Date",
        c(Month, Day, Year),
        sep = "/"
    ) %>%
    head()
```

***
\pagebreak
## 6.4: Data Intake
### Exercise 7

```{r}
url <- "https://en.wikipedia.org/wiki/Mile_run_world_record_progression"
tables <- url %>% read_html() %>% html_nodes("table")
table4 <- rvest::html_table(tables[[4]])

table4 %>% head()
```

***
### Exercise 8

```{r}
url <- "https://en.wikipedia.org/wiki/World_population"
tables <- url %>% 
    rvest::read_html() %>% 
    rvest::html_nodes("table")

rvest::html_table(tables[[5]]) %>% 
    head()
```

***
\pagebreak
## 6.5: Cleaning Data
### Exercise 9

```{r}
# Create houses_small data frame -----
houses_url <-
    "http://sites.msudenver.edu/ngrevsta/wp-content/uploads/sites/416/2021/02/houses-for-sale.txt"

houses <- read.csv(houses_url, header = TRUE, sep = "\t")
houses_small <- select(houses, fuel, heat, sewer, construction)

# Create codebook data frame -----
codebook_url <-
    "http://sites.msudenver.edu/ngrevsta/wp-content/uploads/sites/416/2021/02/house_codes.txt"

translations <- read.csv(codebook_url,
                         header = TRUE,
                         stringsAsFactors = FALSE,
                         sep = "\t")

codes <- translations %>% 
    pivot_wider(names_from = system_type,
                values_from = meaning,
                values_fill = list(meaning = "invalid"))
```

```{r}
dplyr::left_join(
    x = houses_small,
    y = dplyr::select(codes, code, heat_type),
    by = c(heat = "code")
) %>%
    head()
```

***
\pagebreak
### Exercise 10
```{r}
data.frame(Name = c("Joe", "Lucy", "Tom", "Sally"),
                NumberChildren = c("2", "1", "0", "3"),
                stringsAsFactors = FALSE) %>%
    dplyr::mutate(
        NumberChildren = as.numeric(NumberChildren)
    ) %>% 
    str()
```

***
### Exercise 11

```{r}
data.frame(Name = c("Joe", "Lucy", "Tom", "Sally"),
           NumberChildren = c("2", "Unknown", "0", "3"),
           stringsAsFactors = FALSE
          ) %>%
    dplyr::mutate(
        NumberChildren = as.numeric(NumberChildren)
    )
```

The unknown variable in the number children column gets automatically converted to NA. 

***
\pagebreak
### Exercise 12
Guess the output of the following code

Personal note after looking over all of them: I think all of these are equivalent to lubridate, that's my guess after running the first command.

a) lubridate::mdy("Dec 18, 1973")

    - 12/18/1973
    
b) lubridate::mdy("December 18, 1973")

    - "1973-12-18"
    
c) lubridate::mdy("12/18/1973")

    - "1973-12-18"
    
d) lubridate::mdy("12/18/73")

    - "1973-12-18"
    
e) lubridate::mdy("12-18-1973")

    - "1973-12-18"
    
f) lubridate::mdy("12-18-73")

    - "1973-12-18"
    
```{r}
# Checking answers.

glue::glue("
    {lubridate::mdy(\"Dec 18, 1973\")}
    {lubridate::mdy(\"December 18, 1973\")}
    {lubridate::mdy(\"12/18/1973\")}
    {lubridate::mdy(\"12/18/73\")}
    {lubridate::mdy(\"12-18-1973\")}
    {lubridate::mdy(\"12-18-73\")}
")
```

***
\pagebreak
### Exercise 13

```{r}
glue::glue("
    {lubridate::mdy(\"11/14/23\")}
")
```
***
### Exercise 14
How many elapsed days are there between January 15, 20017 and October 4, 2019?

```{r}
day1 <- lubridate::mdy("1/15/07")
day2 <- lubridate::mdy("10/4/19")

glue::glue("
    {day2 - day1} days elapsed between January 15, 2007 and October 4, 2019 
")
```

***
\pagebreak
### Exercise 15
Guess the results of the given commands and check your answers.

a) seq(from = mdy("12-20-1993"), to = mdy("01-15-2004"), by = "days")

    - This will generate a vector of every day between the two provided dates.
    
b) seq(from = mdy("12-20-1993"), to = mdy("01-15-2004"), by = "weeks")

    - This will do the same thing, but instead of every day there will be 7 day gaps.
    
c) seq(from = mdy("12-20-1993"), to = mdy("01-15-2004"), by = "years")

    - This will do the same thing, but with yearly gaps instead of day gaps.
    
```{r}
# A
seq(
    from = lubridate::mdy("12-20-1993"), 
    to = lubridate::mdy("01-15-2004"), 
    by = "days"
) %>% 
    head()
```

```{r}
# B

seq(
    from = lubridate::mdy("12-20-1993"), 
    to = lubridate::mdy("01-15-2004"), 
    by = "weeks"
) %>%
    head()
```

```{r}
# C

seq(
    from = lubridate::mdy("12-20-1993"), 
    to = lubridate::mdy("01-15-2004"), 
    by = "years"
)
```

***
\pagebreak
### Exercise 16

```{r, out.width="60%"}
# A
my.data <- data.frame(Date = c("12/28/2017", "12/29/2017", "12/30/2017",
                               "12/31/2017", "1/1/2018", "1/2/2018", "1/3/2018"),
                      Y = c(44, 43, 47, 53, 53, 55, 56))

ggplot2::ggplot(
    data = my.data,
    mapping = ggplot2::aes(
        x = Date,
        y = Y
    )
) +
    ggplot2::geom_line()
```

It's confused and doesn't know what to make of the date vector. This is unsurprising, they're all strings, not dates. We can easily fix that.

```{r, out.width="60%"}
# B

my.data %>%
    dplyr::mutate(
        Date = lubridate::mdy(Date)
    ) %>%
    ggplot2::ggplot(
        mapping = ggplot2::aes(
            x = Date,
            y = Y
        )
    ) +
    ggplot2::geom_line()
```

***
\pagebreak
# 7: Iteration
## 7.1: Iteration Using for() and best_by() with mutate()
### Exercise 17
My guess is that the below code will print out "good sport" 5 times. I assume this because 1:5 is inclusive, not exclusive. 

```{r}
for (i in 1:5) {
    print("Good Sport")
}
```

***
### Exercise 18
This loop will print each each of the integers in the vector squared. So 4, 16, 36, 64.
```{r}
x <- c(2, 4, 6, 8)

for(i in x) {
    print(i^2)
}
```

### Exercise 19

The sum of squares

\[\sum_{i=1}^{10} i^2 = 1^2 + 2^2 + ... + 10^2\]

can be computed using a for() loop.

```{r}
sum.sq <- 0
for(i in 1:10) {
    sum.sq <- sum.sq + i^2
}
sum.sq
```

a) Why is it necessary to make the assignment sum.sq <- 0 before entering the loop? What would happen if it wasn't there?

    - Assigning it after entering the loop would get it reassigned to 0 at the start of every iteration. If it wasn't there the for loop simply wouldn't work, you'd need to assign something to actually get a meaningful return value, and that assignment would need to happen *outside* of the loop for it to not get immediately overridden. 
    
b) Show what happens if the assignment happens inside the loop.

```{r}
for(i in 1:10) {
    sum.sq <- 0
    sum.sq <- sum.sq + i^2
}
sum.sq
```

***
### Exercise 20

```{r}
(1:10)^2
```

This squares everything in the vector 1:10. You could actually do the same thing as the above (correctly done) sum by getting the sum of this squared vector.

```{r}
(1:10)^2 %>%
    sum()
```

***
\pagebreak
### Exercise 21

```{r}
by_subject <- nest_by(.data = lme4::sleepstudy, Subject)

models <- mutate(.data = by_subject, mod = list(lm(Reaction ~ Days, data = data)))

models %>% head()
```

```{r}
models$mod[[
    which(models$Subject == 371)
]]
```

The equation of the fitted line for subject 371 is:

\[y = 253.636 + 9.188x\]