---
title: "Exercises 9"
author: "Brady Lamson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

### Exercise 1

```{r}
paste("Today's date is", date())
```

```{r}
paste("X", 1:5, sep = "")
```

***
\pagebreak
### Exercise 2

```{r}
# A

first <- "Louis"
middle <- "Daniel"
last <- "Armstrong"

paste(first, middle, last)
```

```{r}
# B

first <- c("Louis", "John", "Miles", "Ella")
middle <- c("Daniel", "William", "Dewey", "Jane")
last <- c("Armstrong", "Coltrane", "Davis", "Fitzgerald")

paste(first, middle, last)
```
```{r}
# C

full <- "Sarah Lois Vaughan"

strsplit(full, split = " ")
```

```{r}
# D

full <- c("Sarah Lois Vaughan", "Thelonious Sphere Monk", "Chet Henry Baker",
"Wynton Learson Marsalis")

strsplit(full, split = " ")
```

***
\pagebreak
### Exercise 3

```{r}
quote <- "Aunt Petunia was horse-faced and bony; Dudley was blond, pink, and
porky. Harry, on the other hand, was small and skinny, with brilliant green
eyes and jet-black hair that was always untidy. He wore round glasses, and on
his forehead was a thin, lightning-shaped scar."

quote <- 
    quote %>%
    gsub(pattern = "\n", replacement = " ") %>%
    gsub(pattern = "[\\.,;]", replacement = "")

quote
```

```{r}
gregexpr(pattern = "-", text = quote)
```
The three values returned represent the indices where dashes appear. That is, the 23rd character, the 149th character and the 246th character. 

```{r}
regexpr(pattern = "-", text = quote)
```

This represents the index of the first instance of a dash appearing in the character string.

***
\pagebreak
### Exercise 4

```{r}
# A

badges <- 
    "Badges? We ain't got no badges. We don't need no badges. I don't have to
show you any stinking badges!" %>%
    gsub(pattern = "\n", replacement = " ") %>%
    tolower()

badges
```

```{r}
# B

badges <-
    badges %>%
    gsub(pattern = "!", replacement = "")

badges
```

```{r}
# C

badges <-
    badges %>%
    gsub(pattern = "[\\?,\\.]", replacement = "")

badges
```

```{r}
# D

badges <-
    badges %>%
    strsplit(split = " ")

badges
```

```{r}
# E

badges <- 
    badges %>%
    unlist()
```

```{r}
badges %>%
    grep(pattern = "badges")

badges %>%
    grepl(pattern = "badges")
```

```{r}
# F

badges %>%
    nchar()
```

***
\pagebreak
### Exercise 5

```{r}
# A / B

quote <- 
    "I have a dream that one day this nation will rise up and live out
the true meaning of its creed, 'We hold these truths to be self-evident,
that all men are created equal.' I have a dream that one day on the
red hills of Georgia, sons of former slaves and the sons of former
slave owners will be able to sit down together at the table of
brotherhood. I have a dream that one day even the state of Mississippi,
a state sweltering with the heat of injustice, sweltering with the
heat of oppression, will be transformed into an oasis of freedom and
justice. I have a dream that my four little children will one day live
in a nation where they will not be judged by the color of their skin
but by the content of their character." %>%
    gsub(pattern = "\n", replacement = " ") %>%
    gsub(pattern = "[\\.,']", replacement = " ") %>%
    strsplit(split = " ") %>%
    unlist()
```

```{r}
q_nchars <- nchar(quote)
```

```{r}
# C

mean(q_nchars)
```

```{r}
# D

ggplot(data = data.frame(n = q_nchars)) +
    geom_histogram(mapping = aes(x = n), fill = "blue", color = "black", binwidth = 1)
```

This histogram is absolutely right-skewed. 

***
\pagebreak
### Exercise 6

```{r}
quote <- "I have a dream that one day this nation will rise up and live out
the true meaning of its creed, 'We hold these truths to be self-evident,
that all men are created equal.' I have a dream that one day on the
red hills of Georgia, sons of former slaves and the sons of former
slave owners will be able to sit down together at the table of
brotherhood. I have a dream that one day even the state of Mississippi,
a state sweltering with the heat of injustice, sweltering with the
heat of oppression, will be transformed into an oasis of freedom and
justice. I have a dream that my four little children will one day live
in a nation where they will not be judged by the color of their skin
but by the content of their character." %>%
    gsub(pattern = "\n", replacement = " ") %>%
    gsub(pattern = "[\\.,']", replacement = "")
```

```{r}
#A 

quote %>%
    gregexpr(pattern = "freedom")
```

The starting character position of "freedom" is index 534. 

```{r}
# B

quote.list <- strsplit(quote, split = " ")
quote.vec <- unlist(quote.list)    # Could also use quote.vec <- quote.list[[1]]
quote.vec
```

This returns everything separated by spaces, which gives us every single individual word as a separate string in a vector.

```{r}
# C

grep(pattern = "the", x = quote.vec)

which(quote.vec == "the")
```

`grep()` simply looks for the letters "the" in each string, the word itself doesn't need to be "the" exactly. The `which()` command is different though, it only returns the indices for the word "the".

```{r}
# D

grep(pattern = "ing$", x = quote.vec, value = TRUE)
```

This checks for strings ending in "ing" specifically. 

```{r}
grep(pattern = "^th", x = quote.vec, value = TRUE)
```

This is the `$` equivalent for the beginning of a string. So this finds all strings that start with "th". 